---
title: "Case Study 1"
author: "Victor Yim"
date: "October 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction


There are many ways to measure and compare economy between countries. The most typical statistics used are either GDP or Income.  
The gross domestic product (GDP) is a unit measure of total values of the final output of the goods and services produced in a year.  The strength of an economy is often measure in GDP and compare year over year to understand growth.  Since this is the total output by a country, there are other variables like population, natural resources, political and other external factors that may impact whether countries have high GDP. 
Average Income is another method to measure strength and prosperity of an economy.  Countries with higher income can typically afford more goods and services, thus in turn create a stronger economy.   As like GDP, there are other factors that could impact this measure.
For this analysis, we attempt to look at the two types for measure to evaluate their relationship.



R codes below explain each of the steps taken to generate the result

```{r cars}
# first setting the directory for the download
setwd("C:/Users/Victor Yim/Documents/SMU/6306/Case Study 1")
# Download data using repmis library
library(repmis)
FinURL1<-"https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv"
GDP <- repmis::source_data(FinURL1, sep = ",", skip=3, header = TRUE)
#Tidy data process
#rename column headers to reflect field values
names(GDP) [1]<-("country")
names(GDP)[5]<-("GDP")
#subsetting data to remove columns with no value
GDP2012<-subset(GDP, country>0, select = c(country, Ranking,Economy,GDP))
#Continue tidy data to reflect actual data characteristics
GDP2012$Ranking<-as.numeric(GDP2012$Ranking)
GDP2012$GDP<-as.numeric(gsub(",","",GDP2012$GDP))
#examine data
head(GDP2012)
#obtain second dataset for income
FinURL2<-"https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
EDU <- repmis::source_data(FinURL2, sep = ",",  header = TRUE)
#rename column variable to match first dataset
names(EDU) [1]<-("country")
#examine data
head(EDU)

```

#Questions on Merged Data
1) Merge the data based on the country shortcode. How many of the IDs match?

```{r}
#merging GDP an EDU dataset from worldbank
total.data <- merge(GDP2012,EDU,by="country")
dim(total.data)
```

There are 224 matches between the 2 datasets

2) Sort the data frame in ascending order by GDP (so United States is last). What is the 13th country in the resulting data frame?



```{r}
total.data2<-subset(total.data,  select = c(country, Ranking,Economy,GDP))
The13th<-total.data2[order(total.data$GDP),]
head(The13th, n=13)
```
there are 2 countries tied for 13:  Grenada and St. Kitts and Nevis


3) What are the average GDP rankings for the "High income: OECD" and "High income:##nonOECD" groups?

```{r}
#tidy data process to remove commas
colnames(total.data) <- gsub(" ","",colnames(total.data))
HIC<-subset(total.data,  select = c(country, Ranking,Economy,GDP,IncomeGroup))
HIC$IncomeGroup<-gsub(" ","",HIC$IncomeGroup)
HIC$IncomeGroup<-gsub(":","",HIC$IncomeGroup)
#subsetting only data with high income
HICOECD<-subset(HIC, IncomeGroup=="HighincomenonOECD" | IncomeGroup=="HighincomeOECD" , NA.RM=TRUE,  select = c(GDP, country, IncomeGroup))
aggregate( GDP~IncomeGroup, HICOECD, mean )

```
The Average GDP for HighincomenonOECD = 104349.8 and the Average GDP for HighincomeOECD  = 1483917.1

4) Plot the GDP for all of the countries. Use ggplot2 to color your plot by Income Group.

```{r}
library(ggplot2)
HICGRAPH<-subset(HIC, NA.RM=TRUE, GDP>1,  select = c(GDP, country, IncomeGroup))
HICGRAPH2<-HICGRAPH[complete.cases(HICGRAPH),]
#qplot(x, y, data=, color=, shape=, size=, alpha=, geom=, method=, formula=, facets=, xlim=, ylim= xlab=, ylab=, main=, sub=)
qplot(country, GDP, data=HICGRAPH2, color=factor(IncomeGroup),  main='GDP by Country')

```

5) Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group.
How many countries are Lower middle income but among the 38 nations with highest
GDP?

```{r}
library(dplyr)
HIC$quantile <- ntile(HIC$GDP, 5) 
attach(HIC)
Pivottabletable <- table(IncomeGroup, quantile)
detach(HIC)

Pivottabletable
```

There are 17 countries that are Lower middle income but among the 38 nations with highest GDP (in first quantity above)


## Summary
After analyzing the GDP data from Worldbank, there appears to be a significant range between the countries' gross domestic product output. the top 20% of the countries has an average GDP > 1000 times greater than the botton 20% of the countries.
By using the Organization for Economic Co-operation and Development (OECD) comparison,  we  found that countries with OECD membership has a higher average GDP than those that are not members.
Using the 5 income tier provided by the worldbank:   
- High income: nonOECD  
- High income: OECD  
- Lower middle income  
- Upper middle income  
- Low income  
we found that the average income level does not also result in higher GDP or vise versa.   These measures are useful to answer specific questions in related to the topics.  

